---
title: "ISL_CW_24"
author: "Luyuan Wang"
date: "2024-02-18"
output: html_document
---

happiness index.
Reasons for the choice of data:
Personal Interest: 
Reference for policy-making decisions for governments, organizations and civil society.
proclaimed on 3.20 of the year to be aligned with the International Day of Happiness.

```{r data preparation}
setwd("C:/Users/Luyuan Wang/Documents/R")
H_index <- read.csv("WHR2023.csv")

row.names(H_index) <- H_index[,1]

# omit rows with NA
new_H_index <- na.omit(H_index)
```

```{r}
# read the Asian and European country list and select from the H_index data frame
setwd("C:/Users/Luyuan Wang/Documents/R")
asia_euro <- read.csv("asia_euro.csv")
c_list <- asia_euro[,1]
c_list <- c_list[duplicated(c_list)==FALSE]
# row.names(H_index) <- H_index[,1]
new_H_index <- H_index[c_list, ]

# the index for Asia
asia <- read.csv("asia.csv")
a_list <- asia[,1]
a_list <- a_list[duplicated(a_list)==FALSE]
a_H_index <- new_H_index[a_list, ]

# the index for Europe
eu <- read.csv("europe.csv")
eu_list <- eu[,1]
eu_list <- eu_list[duplicated(eu_list)==FALSE]
eu_H_index <- new_H_index[eu_list, ]
```

```{r}
new_H_index <- na.omit(new_H_index)
config_x <- new_H_index[, 2:8]
config_x <- na.omit(config_x)
```

```{r implement mdscale}
# as distance object
dist_H <- dist(config_x[, 2:7], method="euclidean")

# perform mds
fit_H <- cmdscale(dist_H, eig=TRUE, k=2)

# criteria of judging the number of dimensions at: https://onlinelibrary.wiley.com/doi/epdf/10.1002/bimj.4710240520?saml_referrer
eigs <- fit_H$eig
print(sum((eigs[1:2])^2)/sum((eigs)^2))  # 0.999996

#extract (x, y) coordinates of multidimensional scaling
x <- fit_H$points[,1]
y <- fit_H$points[,2]
```
Observation: where elements of groups have overlapping names
visualize their ladder score by groups
indeed, similar scores with range (TBA)

```{r plots}
#create scatter plot
plot(x, y, xlab="Coordinate 1", ylab="Coordinate 2",
     main="Multidimensional Scaling Results of Ladder Score", type="n")

#add row names of data frame as labels
text(x, y, labels=row.names(config_x), cex=0.4)

#create scatter plot using ladder score
plot(x, y, xlab="Coordinate 1", ylab="Coordinate 2",
     main="Strtified Ladder Score Distribution", type="n")
text(x, y, labels=new_H_index[, "Ladder.score"], cex=0.4)

# borders
abline(a=-1.55, b=-3/10, lty=2, col=1)
abline(a=-0.1, b=-7/40, lty=2, col=2)
abline(a=0.9, b=-7/40, lty=2, col=3)
abline(a=2, b=-3/20, lty=2, col=4)
abline(a=4, b=-1/5, lty=2, col=5)
```


```{r}
# in terms of closeness, hand-pick
```

```{r, visualizing ladder scores of overlapping areas}

library("ggplot2")

# construct dataframe for ggplot
l_score <- as.data.frame(cbind(new_H_index[,1], config_x[,"Ladder.score"]))
colnames(l_score) <- c("Country.name", "Ladder.score")

g_1 <- group(x_cor=c(-5, -1.5), y_cor=c(-0.5, -0.35), x, y)
g1_p <- ggplot(data = l_score[g_1, ], aes(x = Country.name, y = Ladder.score)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = Ladder.score), vjust=1.6, 
            color="black", size=2.5)

# Horizontal bar plot
g1_p + coord_flip()

# europe
g_2 <- group(x_cor=c(-1, 1), y_cor=c(0.3, 0.5), x, y)
g2_p <- ggplot(data = l_score[g_1, ], aes(x = Country.name, y = Ladder.score)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = Ladder.score), vjust=1.6, 
            color="black", size=2.5)

# Horizontal bar plot
g2_p + coord_flip()


g_3 <- group(x_cor=c(-3, 0), y_cor=c(-0.1, 0.05), x, y)
g3_p <- ggplot(data = l_score[g_1, ], aes(x = Country.name, y = Ladder.score)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = Ladder.score), vjust=1.6, 
            color="black", size=2.5)

# Horizontal bar plot
g3_p + coord_flip()

```


Notice the overall score of the region 1 is higher than those of region 2 and 3. Noticing that from Europe and Asia respectively, visualized the average scores across continents.

What makes European happier.
```{r ,average of europe compared with that of asia}

avr_a <- round(mean(a_H_index[,"Ladder.score"], na.rm=TRUE), 2) 
avr_eu <- round(mean(eu_H_index[,"Ladder.score"]), 2)

barplot(c(avr_a, avr_eu), names.arg=c("Asia","Europe"), col="blue", xlab="Continent", ylab="Average Ladder Score", main="The average ladder score of Asia and Europe in 2023", ylim=c(0,7))
text(x=c(0.7,1.9), y=c(5.15,6.25), label=c(avr_a, avr_eu), pos = 3, cex = 0.8, col = "black")
```
## Grouping functions
```{r function to extract the elemts}
group <- function(x_cor, y_cor, x, y){
  x_index <- c()
  index_vec <- c()
  for (i in 1:length(x)) {
    if(x[i] > x_cor[1] & x[i] < x_cor[2]){
     x_index <- append(x_index, i)
    }
  }
  for (i in x_index){
    if(y[i] > y_cor[1] & y[i] < y_cor[2]){
      index_vec <- append(index_vec, i)
    }
  }
  return(index_vec)
}
```

```{r}

# g_1
abline(v=-5, lty=2, col=3)
abline(v=-1.5, lty=2, col=3)
abline(h=-0.5, lty=2, col=3)
abline(h=-0.35, lty=2, col=3)

# g_2
abline(v=-1, lty=2, col=2)
abline(v=1, lty=2, col=2)
abline(h=0.3, lty=2, col=2)
abline(h=0.5, lty=2, col=2)

# g_3
abline(v=-3, lty=2, col=4)
abline(v=0, lty=2, col=4)
abline(h=-0.1, lty=2, col=4)
abline(h=0.05, lty=2, col=4)
```

```{r}
# omit rows with NA
new_H_index <- na.omit(new_H_index)

# delete the row of Palestine
# H_index <- H_index[-c(H_index[,1]=='State of Palestine'),]
```

```{r plots of eigenvalues of fit_H}
eigs <- fit_H$eig
plot(1:length(eigs), log(abs(eigs)))
points(1:37, log(abs(eigs[1:37])), col=1)
points(38: length(eigs), log(abs(eigs[38:length(eigs)])), col=4)
```
